A patch-based framework for new ITK functionality: Joint fusion, denoising, and non-local super-resolution

### 2017-03-15 Document/LaTeX/ArticleExample.tex
ENH:  Add author.

% At minimum, give your name and an email address.  You can include a
% snail-mail address if you like.
\author{Nick Tustison$^{1}$ Brian Avants$^{2}$, Hongzhi Wang$^{3}$, Long Xie$^{4}$, Paul Yushkevich$^{4}$, and Jose Vicente Manj\'on$^{5}$}
\authoraddress{$^{1}$University of Virginia, Charlottesville, VA\\
               $^{2}$Biogen, Cambridge, MA\\
               $^{3}$IBM, San Jose, CA\\
               $^{4}$University of Pennsylviania, Philadelphia, PA\\
               $^{5}$ITACA institute, Universidad Polit\'ecnica de Valencia, Valencia, Spain
               }

\begin{document}

@@

Since the non-local (NL) means algorithm introduced in several early papers \cite{Buades},
the resulting family of research techniques has grown to significant prominence within the
medical imaging community.  In addition to continued research output, patch-based techniques
have been the subject of workshops (Patch-MI 2015 and PMI'16, both MICCAI-sponsored
workshops).   Given the interest and potential applications, this submission is a potential
``conversation starter'' for the development of an ITK software framework

@@

The following files define the adaptive NL-means algorithm and is written in ITK
coding style:
\begin{itemize}
  \item itkAdaptiveNonLocalMeansDenoisingImageFilter.h
  \item itkAdaptiveNonLocalMeansDenoisingImageFilter.hxx
\end{itemize}
The following helper class
\begin{itemize}
  \item itkVarianceImageFilter.h
  \item itkVarianceImageFilter.hxx
\end{itemize}
is the variance analog to the \doxygen{MeanImageFilter} which ``Computes an image where a
given pixel is the [variance] value of the the pixels in a neighborhood about the corresponding
input pixel.''  Both the adaptive and variance filters are multi-threaded.

Additionally, we wrote an easy-to-use interface program, DenoiseImage.cxx, with an ANTs-based command
line interface facilitated by the following helper classes:
\begin{itemize}
  \item antsCommandLineOption.h
  \item antsCommandLineOption.cxx
  \item antsCommandLineParser.h
  \item antsCommandLineParser.cxx
\end{itemize}
These latter files are not essential for other applications or for inclusion into ITK.
They are included simply to facilitate the generation of a stand-alone application that
people can use.  One can invoke the help menu via \verb#$ DenoiseImage --help 1# at the
command line:
\small
\begin{verbatim}
$ DenoiseImage --help 1

@@

     --help
          Print the help menu.
\end{verbatim}
\normalsize
which should provide minimal detail regarding each parameter.  The default parameters
should provide adequate performance for a variety of cases thus requiring only the simple
call (for a 3-D image)

\small
\begin{verbatim}
$ DenoiseImage -d 3 -i ${inputImage} -o ${outputImage}
\end{verbatim}
\normalsize

As mentioned, this filter is multi-threaded so one might want to set the environment
variable to enable more than one thread, e.g.,
\small
\begin{verbatim}
$ export ITK_GLOBAL_DEFAULT_NUMBER_OF_THREADS=4
\end{verbatim}
\normalsize

To see the progress of the filter, one also needs to set the verbose flag, i.e.,
\verb#-v 1#.

\section{Examples}

We included a 2-D brain slice and a 3-D hyperpolarized gas volume each with a corresponding
mask.  Below we include the call and resulting output for these two sets of data.

\subsection{T1-weighted MRI brain slice}

\small
\begin{verbatim}
$ DenoiseImage -d 2 -i t81slice.nii.gz -x t81mask.nii.gz -o t81denoised.nii.gz -v 1

Running for 2-dimensional images.

*********10*********20*********30*********40*********50*********60*********70*********80*********90*********

   AdaptiveNonLocalMeansDenoisingImageFilter (0x7f9733997080)
     RTTI typeinfo:   itk::AdaptiveNonLocalMeansDenoisingImageFilter<itk::Image<float, 2u>, itk::Image<float, 2u>, itk::Image<unsigned char, 2u> >
     Reference Count: 1
     Modified Time: 517
     Debug: Off
     Object Name:
     Observers:
       ProgressEvent(Command)
     Inputs:
       Primary: (0x7f9733996c10) *
       _1: (0x7f97339988f0)
     Indexed Inputs:
       0: Primary (0x7f9733996c10)
       1: _1 (0x7f97339988f0)
     Required Input Names: Primary
     NumberOfRequiredInputs: 1
     Outputs:
       Primary: (0x7f9733996a50)
     Indexed Outputs:
       0: Primary (0x7f9733996a50)
     NumberOfRequiredOutputs: 1
     Number Of Threads: 1
     ReleaseDataFlag: Off
     ReleaseDataBeforeUpdateFlag: Off
     AbortGenerateData: Off
     Progress: 1
     Multithreader:
       RTTI typeinfo:   itk::MultiThreader
       Reference Count: 1
       Modified Time: 299
       Debug: Off
       Object Name:
       Observers:
         none
       Thread Count: 1
       Global Maximum Number Of Threads: 128
       Global Default Number Of Threads: 1
     CoordinateTolerance: 1e-06
     DirectionTolerance: 1e-06
     Using Gaussian noise model.
     Epsilon = 1e-05
     Mean threshold = 0.95
     Variance threshold = 0.5
     Smoothing variance = 2
     Neighborhood radius for local mean and variance = [1, 1]
     Neighborhood search radius  = [3, 3]
     Neighborhood block radius = [1, 1]
Elapsed time: 0.0831451
\end{verbatim}
\normalsize

\begin{figure}[!ht]
\begin{subfigure}{0.5\textwidth}
  \centering
  \includegraphics[width=0.8\linewidth]{Figures/t81slice.eps}
  \caption{Original}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
  \centering
  \includegraphics[width=0.8\linewidth]{Figures/t81denoised.eps}
  \caption{Denoised}
\end{subfigure}
\caption{Effects of the NL-means filter on a 2-D T1-weighted MRI brain slice.}
\end{figure}

\subsection{Hyperpolarized He3 Lung MRI volume}

\small
\begin{verbatim}
$ DenoiseImage -d 3 -i he3volume.nii.gz -x he3mask.nii.gz -o he3denoised.nii.gz -v 1

Running for 3-dimensional images.

*********10*********20*********30*********40*********50*********60*********70*********80*********90*********

   AdaptiveNonLocalMeansDenoisingImageFilter (0x7ff279e0a440)
     RTTI typeinfo:   itk::AdaptiveNonLocalMeansDenoisingImageFilter<itk::Image<float, 3u>, itk::Image<float, 3u>, itk::Image<unsigned char, 3u> >
     Reference Count: 1
     Modified Time: 506
     Debug: Off
     Object Name:
     Observers:
       ProgressEvent(Command)
     Inputs:
       Primary: (0x7ff27a97ea20) *
       _1: (0x7ff27a97f040)
     Indexed Inputs:
       0: Primary (0x7ff27a97ea20)
       1: _1 (0x7ff27a97f040)
     Required Input Names: Primary
     NumberOfRequiredInputs: 1
     Outputs:
       Primary: (0x7ff279e0a710)
     Indexed Outputs:
       0: Primary (0x7ff279e0a710)
     NumberOfRequiredOutputs: 1
     Number Of Threads: 4
     ReleaseDataFlag: Off
     ReleaseDataBeforeUpdateFlag: Off
     AbortGenerateData: Off
     Progress: 1
     Multithreader:
       RTTI typeinfo:   itk::MultiThreader
       Reference Count: 1
       Modified Time: 294
       Debug: Off
       Object Name:
       Observers:
         none
       Thread Count: 4
       Global Maximum Number Of Threads: 128
       Global Default Number Of Threads: 4
     CoordinateTolerance: 1e-06
     DirectionTolerance: 1e-06
     Using Gaussian noise model.
     Epsilon = 1e-05
     Mean threshold = 0.95
     Variance threshold = 0.5
     Smoothing variance = 2
     Neighborhood radius for local mean and variance = [1, 1, 1]
     Neighborhood search radius  = [3, 3, 3]
     Neighborhood block radius = [1, 1, 1]
Elapsed time: 1.7134
\end{verbatim}
\normalsize

\begin{figure}[!ht]
\begin{subfigure}{0.5\textwidth}
  \centering
  \includegraphics[width=0.8\linewidth]{Figures/he3volume.eps}
  \caption{Original}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
  \centering
  \includegraphics[width=0.8\linewidth]{Figures/he3denoised.eps}
  \caption{Denoised}
\end{subfigure}
\caption{Effects of the NL-means filter on a 3-D hyperpolarized gas MRI lung volume.}
\end{figure}

>>>>

% At minimum, give your name and an email address.  You can include a
% snail-mail address if you like.
\author{Nick Tustison$^{1}$
        Brian Avants$^{2}$,
        Hongzhi Wang$^{3}$,
        Long Xie$^{5}$,
        Pierrick Coup\'e$^{4}$,
        Paul Yushkevich$^{5}$, and
        Jose Vicente Manj\'on$^{5}$}
\authoraddress{$^{1}$University of Virginia, Charlottesville, VA, USA\\
               $^{2}$Biogen, Cambridge, MA, USA\\
               $^{3}$IBM Research, San Jose, CA, USA\\
               $^{4}$ Laboratoire Bordelais de Recherche en Informatique, Talence, France\\
               $^{5}$University of Pennsylviania, Philadelphia, USA\\
               $^{6}$ITACA institute, Universidad Polit\'ecnica de Valencia, Valencia, Spain
               }
\begin{document}

@@

Since the non-local (NL) means algorithm introduced in several early papers,
the resulting family of research techniques has grown to significant prominence within the
medical imaging community.  In addition to continued research output, patch-based techniques
have been the subject of workshops (Patch-MI 2015 and PMI'16, both MICCAI-sponsored
workshops).   Given the interest and potential applications, this submission is a potential
``conversation starter'' for the development of an ITK software framework for accommodating
such techniques.

@@

\subsection{Parent and child classes}

Based on the implementations of different algorithms from the various authors,
the common routines were grouped to the following parent class:

\begin{itemize}
  \item itkNonLocalPatchBasedImageFilter.h
  \item itkNonLocalPatchBasedImageFilter.hxx
\end{itemize}

The common routines contained within this class include defining both the
patch and search neighborhoods.  We also define two patch similarity
metrics---sum of squared differences (SSD) and Pearson's correlation.
We also provide methods for ``vectorizing'' image patches and multiple image
patches (for multi-modal applications) for facilitating similarity metric
calculation.

We provide the following classes for the adaptive denoising algorithm:
\begin{itemize}
  \item itkAdaptiveNonLocalMeansDenoisingImageFilter.h
  \item itkAdaptiveNonLocalMeansDenoisingImageFilter.hxx
\end{itemize}
The following helper class
\begin{itemize}
  \item itkVarianceImageFilter.h
  \item itkVarianceImageFilter.hxx
\end{itemize}
These have been refactored since our previous contribution such that the
adaptive denoising class is derived from the parent class.  Again, the
second class is the variance analog to the \doxygen{MeanImageFilter}
which ``Computes an image where a
given pixel is the [variance] value of the the pixels in a neighborhood about the corresponding
input pixel.''

For the super-resolution algorithm, we provide the following class:
\begin{itemize}
  \item itkNonLocalSuperresolutionImageFilter.h
  \item itkNonLocalSuperresolutionImageFilter.hxx
\end{itemize}
which subsumes functionalities described in both \cite{Manjon:2010ab,Manjon:2010ac}.

Finally, for the joint fusion algorithm \cite{Wang:2013ab,Wang:2013aa} we provide
the following class:
\begin{itemize}
  \item itkWeightedVotingFusionImageFilter.h
  \item itkWeightedVotingFusionImageFilter.hxx
\end{itemize}

Note that all classes are multi-threaded.

\subsection{Programs}

We provide the following routines which use the above-listed classes:
\begin{itemize}
  \item DenoiseImage.cxx
  \item NonLocalSuperResolution.cxx
  \item jointFusion.cxx
\end{itemize}

All these routines use an ANTs-based command line interface facilitated by the
following helper classes:
\begin{itemize}
  \item antsCommandLineOption.h
  \item antsCommandLineOption.cxx
  \item antsCommandLineParser.h
  \item antsCommandLineParser.cxx
\end{itemize}
These latter files are not essential for other applications or for inclusion into ITK.
They are included simply to facilitate the generation of a stand-alone application that
people can use.

One can invoke the help menu for all these routines via the \verb#--help# option invoked at the
command line.  To see the progress of the filter, one also needs to set the verbose flag, i.e.,
\verb#-v 1#.  Below we give the help menu for each program.

\subsection{Denoise image}

\small
\begin{verbatim}
$ DenoiseImage --help 1

@@

     --help
          Print the help menu.
\end{verbatim}
\normalsize

\subsection{Super-resolution}

\small
\begin{verbatim}
$ NonLocalSuperResolution --help 1

COMMAND:
     NonLocalSuperResolution
          Non-local super resolution described in the following papers: 1) JV Manjon, P
          Coupe, A Buades, V Fonov, DL Collins, and Montserrat Robles. Non-local MRI
          Upsampling.Medical Image Analysis, 14:784-792, 2010 and2) JV Manjon, P Coupe, A
          Buades, DL Collins, and Montserrat Robles. MRI Superresolution Using
          Self-Similarity and Image Priors.International Journal of Biomedical Imaging,
          2010.

     -i, --input-image inputImageFilename
          A low-resolution image input image to be superresoluted.

     -j, --interpolated-image inputImageFilename
          An interpolated version of the low-resolution image (such as B-spline). One
          should specify either this option as a secondary input or a high-resolution
          multi-modal counterpart (cf the -k option).

     -k, --reference-image inputImageFilename
          A high resolution reference multi-modal image. Assumed to be in the same space
          as the low-resolution input image (i.e., registered).One should specify either
          this option as a secondary input or an interpolated version (cf the -j option).

     -g, --intensity-difference-sigma 1.0
          Intensity difference sigma. Default = 1.0

     -t, --patch-similarity-sigma 1.0
          Patch similarity sigma. Default = 1.0

     -s, --scale-levels 32x16x8x2x1
          Scale levels. Default = 32x16x8x2x1

     -n, --interpolation Linear
                         NearestNeighbor
                         Gaussian[<sigma=imageSpacing>,<alpha=1.0>]
                         BSpline[<order=3>]
                         CosineWindowedSinc
                         WelchWindowedSinc
                         HammingWindowedSinc
                         LanczosWindowedSinc
          Several interpolation options are available in ITK. These have all been made
          available.

     -o, --output outputImage
          The output consists of the noise corrected version of the input image.
          Optionally, one can also output the estimated noise image.

     --version
          Get Version Information.

     --help
          Print the help menu.
          <VALUES>: 1
\end{verbatim}
\normalsize

\subsection{Joint fusion}

\small
\begin{verbatim}
$ jointFusion --help 1

COMMAND:
     antsJointFusion
          antsJointFusion is an image fusion algorithm developed by Hongzhi Wang and Paul
          Yushkevich which won segmentation challenges at MICCAI 2012 and MICCAI 2013. The
          original label fusion framework was extended to accommodate intensities by Brian
          Avants. This implementation is based on Paul's original ITK-style implementation
          and Brian's ANTsR implementation. References include 1) H. Wang, J. W. Suh, S.
          Das, J. Pluta, C. Craige, P. Yushkevich, Multi-atlas segmentation with joint
          label fusion IEEE Trans. on Pattern Analysis and Machine Intelligence, 35(3),
          611-623, 2013. and 2) H. Wang and P. A. Yushkevich, Multi-atlas segmentation
          with joint label fusion and corrective learning--an open source implementation,
          Front. Neuroinform., 2013.

     -t, --target-image targetImage
                        [targetImageModality0,targetImageModality1,...,targetImageModalityN]
          The target image (or multimodal target images) assumed to be aligned to a common
          image domain.

     -g, --atlas-image atlasImage
                       [atlasImageModality0,atlasImageModality1,...,atlasImageModalityN]
          The atlas image (or multimodal atlas images) assumed to be aligned to a common
          image domain.

     -l, --atlas-segmentation atlasSegmentation
          The atlas segmentation images. For performing label fusion the number of
          specified segmentations should be identical to the number of atlas image sets.

     -a, --alpha 0.1
          Regularization term added to matrix Mx for calculating the inverse. Default =
          0.1

     -b, --beta 2.0
          Exponent for mapping intensity difference to the joint error. Default = 2.0

     -c, --constrain-nonnegative (0)/1
          Constrain solution to non-negative weights.

     -p, --patch-radius 2
                        2x2x2
          Patch radius for similarity measures. Default = 2x2x2

     -m, --patch-metric (PC)/MSQ
          Metric to be used in determining the most similar neighborhood patch. Options
          include Pearson's correlation (PC) and mean squares (MSQ). Default = PC (Pearson
          correlation).

     -s, --search-radius 3
                         3x3x3
                         searchRadiusMap.nii.gz
          Search radius for similarity measures. Default = 3x3x3. One can also specify an
          image where the value at the voxel specifies the isotropic search radius at that
          voxel.

     -e, --exclusion-image label[exclusionImage]
          Specify an exclusion region for the given label.

     -x, --mask-image maskImageFilename
          If a mask image is specified, fusion is only performed in the mask region.

     -o, --output labelFusionImage
                  intensityFusionImageFileNameFormat
                  [labelFusionImage,intensityFusionImageFileNameFormat,<labelPosteriorProbabilityImageFileNameFormat>,<atlasVotingWeightImageFileNameFormat>]
          The output is the intensity and/or label fusion image. Additional optional
          outputs include the label posterior probability images and the atlas voting
          weight images.

     --version
          Get version information.

     --help
          Print the help menu.
          <VALUES>: 1
\end{verbatim}
\normalsize

\section{Usage examples}

Instead of providing examples contained within the submission, we point to
self-contained github repos which showcase usage for each of the three
programs.

\subsection{Denoise image}

\begin{itemize}
  \item \url{https://github.com/ntustison/DenoiseImageExample}
\end{itemize}

\subsection{Super-resolution}

\begin{itemize}
  \item \url{https://github.com/ntustison/NonLocalSuperResolutionExample}
\end{itemize}

\subsection{Joint fusion}

Note that for the following example, the user will need to replace ``antsJointFusion''
with ``jointFusion'' due to some legacy issues in preparing this submission.

\begin{itemize}
  \item \url{https://github.com/ntustison/MalfLabelingExample}
\end{itemize}
